---
title: "README"
author: "Jerry"
date: "2024-12-17"
output: pdf_document
---

# Project Title



## Description

This study investigates the relationship between carbon dioxide (CO$_2$) emissions and farmland usage across the contiguous United States from 1970 to 2012, recorded at 5-year intervals. By analyzing state-level data on CO$_2$ emissions and farmland acreage, we aim to determine whether states with a larger percentage of farmland produce lower CO$_2$ emissions per unit area.

## Getting Started

### Dependencies
Library needed for data processing: 
```{r preload, results=FALSE, message=FALSE, warning=FALSE}
# if you do not have USAboundaries installed:
install.packages("remotes")
install.packages("sf")
install.packages("lmtest")
library(lmtest)
library(remotes)
library(tidyverse)
library(readxl)
library(sf)
library(dplyr)
library(plm)

remotes::install_github("ropensci/USAboundaries")
```


All data can be found in (https://github.com/chenyun0607/STAT615), please make sure to download and process the following codes.



### Executing program
Below are the codes that started with cleaning&pivoting data 
```{r clean-up, results=FALSE, message=FALSE, warning=FALSE}
# Importing data on co2 emissions by state
co2_emissions_state <- read_excel("carbon_emissions_by_state.xlsx", skip=4) %>%
  head(-2) %>% filter(!(State %in% c("Alaska", "Hawaii", "District of Columbia")))

# Importing data on cropland size by state + regions
cropland_state_regions <- read_excel("cropland_by_state.xls", skip=2, na="-") %>% 
  na.omit() %>% 
  rename("RegionsAndStates" = "Regions and States", "2012" = "2012 5/")

# Importing data on total land size by state + regions
land_area_state_regions <- read_excel("total_land_by_state.xls", skip=2, na="-") %>% 
  na.omit() %>% 
  rename("RegionsAndStates" = "Regions and States")

# Fixing some column types error:
cropland_state_regions$'1945' <- as.numeric(cropland_state_regions$'1945')
cropland_state_regions$'1969' <- as.numeric(cropland_state_regions$'1969')
land_area_state_regions$'1969' <- as.numeric(land_area_state_regions$'1969')


```

```{r pivot, results=FALSE, message=FALSE, warning=FALSE}
# Pivoting all imported data:
cropland_pivoted <- cropland_state_regions %>% 
  pivot_longer(cols = -RegionsAndStates, names_to = "Year", values_to = "Cropland_Area")

land_area_pivoted <- land_area_state_regions %>% 
  pivot_longer(cols = -RegionsAndStates, names_to = "Year", values_to = "Total_Land_Area")

co2_emissions_pivoted <- co2_emissions_state %>% 
  pivot_longer(cols = -State, names_to = "Year", values_to = "Total_Emissions")

co2_emissions_pivoted <- left_join(co2_emissions_pivoted,
                                   land_area_pivoted,
                                   by=c("State"="RegionsAndStates","Year")) %>% 
  mutate(percent_co2 = (Total_Emissions / Total_Land_Area) * 100) %>% 
  drop_na()


  
excluded_regions <- c("Appalachian", "Corn Belt", "Delta States", 
                      "Lake States", "Mountain", "Northeast", 
                      "Northern Plains", "Southeast", "Southern Plains", "Pacific", "District of Columbia 2/", "48 States 1/", "Alaska 3/", "Hawaii 4/", "U.S. Total 1/")


cropland_pivoted <- cropland_pivoted%>%filter(Year >= 1974)%>%filter(!is.na(Cropland_Area))%>% 
  filter(!RegionsAndStates %in% excluded_regions)

cropland_pivoted_sorted <- cropland_pivoted[order(cropland_pivoted$RegionsAndStates), ]
cropland_pivoted$RegionsAndStates <- factor(cropland_pivoted$RegionsAndStates, levels = sort(unique(cropland_pivoted$RegionsAndStates)))

df <- cbind(cropland_pivoted, co2_emission = co2_emissions_pivoted$Total_Emissions)
```


Below are codes used for analysis:
```{r regression , results=FALSE, message=FALSE, warning=FALSE}

# Ensure your data is ordered by state and year
df <- df %>%
  arrange(RegionsAndStates, Year)

# Convert your dataset into a panel data frame
# "RegionsAndStates" identifies the cross-sectional unit (state)
# "Year" identifies the time dimension

pdata <- pdata.frame(df, index = c("RegionsAndStates", "Year"))

# Run a panel regression
# Here is an example model where we try to explain Cropland_Area by co2_emission.

model <- plm(co2_emission ~ Cropland_Area,
             data = pdata,
             model = "within",
             effect = "twoways") # "within" = fixed effects model


summary(model)

linear_model <- lm(co2_emission ~ poly(Cropland_Area, 2),
           data = pdata)



#testing normality assumptions to see if the regular linear model is usable
#test constant variance: no good 
bptest(linear_model)
#test residual normality: no good 
shapiro.test(residuals(linear_model))
#introduce lagged lm
df$co2_lag <- dplyr::lag(df$co2_emission, 1)
model_lag <- lm(co2_emission ~ poly(Cropland_Area, 2) + co2_lag, data = df)
model_lag
#test autocorrelation: no good 
dwtest(model_lag)


#so using non-parametric method: Bootstrapping
set.seed(123)  # For reproducibility

# Fit the original model
original_model <- lm(co2_emission ~ poly(Cropland_Area, 2), data = df)

# Extract the original coefficients
original_coefs <- coef(original_model)

# Set the number of bootstrap iterations
B <- 1000  # You can increase to 2000 or more for better accuracy

# Create a storage matrix for bootstrapped coefficients
boot_coefs <- matrix(NA, nrow = B, ncol = length(original_coefs))

# Perform the bootstrap
for (b in 1:B) {
  # Sample indices with replacement
  boot_indices <- sample(seq_len(nrow(df)), size = nrow(df), replace = TRUE)
  boot_sample <- df[boot_indices, ]
  
  # Refit the model on the bootstrap sample
  boot_model <- lm(co2_emission ~ poly(Cropland_Area, 2), data = boot_sample)
  
  # Store the coefficients
  boot_coefs[b, ] <- coef(boot_model)
}

# Compute bootstrap standard errors
boot_se <- apply(boot_coefs, 2, sd)

# Get bootstrap confidence intervals (e.g., 95%)
alpha <- 0.05
boot_ci <- apply(boot_coefs, 2, quantile, probs = c(alpha/2, 1 - alpha/2))

# Summarize results
results <- data.frame(
  Estimate = original_coefs,
  Boot_SE = boot_se,
  CI_lower = boot_ci[1, ],
  CI_upper = boot_ci[2, ]
)

print(results)
```


## Author
Jerry, Chen-Yun, Yang


